---
title: "Step 4.1 Merge all samples by Seurat and Python - scVI"
author: "Zhang Yongchao"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: TRUE
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      warning = FALSE)
```

Method Refer:

>[Seurat](https://satijalab.org/seurat/)


>[Tutorial: Integrating stimulated vs. control PBMC datasets to learn cell-type specific responses](https://satijalab.org/seurat/archive/v3.1/immune_alignment.html)

**Parameters:**

```{r}
Results.dir <- "./4.1-Merged-Rds/scVI-Method"
cluster.resolution <- "CellBender_snn_res.0.6"
```


```{r}
suppressMessages(library(dplyr)) 
suppressMessages(library(Seurat))
suppressMessages(library(patchwork))
suppressMessages(library(cowplot))
suppressMessages(library(ggplot2))
suppressMessages(library(ggpubr))
suppressMessages(library(clustree))
suppressMessages(library(harmony))
suppressMessages(library(scater))
suppressMessages(library(DT))
suppressMessages(library(reticulate))
```

# import and merge all samples

```{r}
immune.combined <- readRDS("./4.1-Merged-Rds/Merge-All-SeuratHarmony.rds")
```

# Modify old cluster names - Prefixed with Single

```{r}
colnames(immune.combined@meta.data)[grepl("(S.Score)|(G2M.Score)|(Phase)|(_snn_res)|(^Annotation)",colnames(immune.combined@meta.data))] <- 
  paste("Merge-Harmony_",colnames(immune.combined@meta.data)[grepl("(S.Score)|(G2M.Score)|(Phase)|(_snn_res)|(^Annotation)",colnames(immune.combined@meta.data))],sep = "")
```

## Perform an integrated analysis

```{r}
sc <- import('scanpy', convert = FALSE)
scvi <- import('scvi', convert = FALSE)
scvi$settings$progress_bar_style = 'tqdm'
```

```{r}
as_matrix <- function(mat){

  tmp <- matrix(data=0L, nrow = mat@Dim[1], ncol = mat@Dim[2])
  
  row_pos <- mat@i+1
  col_pos <- findInterval(seq(mat@x)-1,mat@p[-1])+1
  val <- mat@x
    
  for (i in seq_along(val)){
      tmp[row_pos[i],col_pos[i]] <- val[i]
  }
    
  row.names(tmp) <- mat@Dimnames[[1]]
  colnames(tmp) <- mat@Dimnames[[2]]
  return(tmp)
}
x <- as_matrix(GetAssayData(immune.combined,slot='counts'))
# x <- x[VariableFeatures(immune.combined),]
obs <- immune.combined[[]]
var <- GetAssay(immune.combined)[[]]
# var <- var[VariableFeatures(immune.combined),]
dim(var)
```



```{r}
adata <- sc$AnnData(
  X   = t(x), #scVI requires raw counts
  obs = obs,
  var = var
)
print(adata) # Note generally in Python, dataset conventions are obs x var
```


```{r}
# run seteup_anndata
scvi$data$setup_anndata(adata,
                        # layer="counts",
                        batch_key = "orig.ident",
                        # categorical_covariate_keys = c("subject","sex"),
                        continuous_covariate_keys = c("nCount_RNA", "percent.mt","Merge-Harmony_S.Score", "Merge-Harmony_G2M.Score"))

# create the model
model = scvi$model$SCVI(adata,
                        n_hidden=as.integer(128),  # Attention, use as.interger for numbers, otherwise numbers will be recognized as folat type in python.
                        n_latent=as.integer(20), 
                        n_layers=as.integer(2))
# run this step need about 6 hours


# train the model
model$train()

# to specify the number of epochs when training:
# model$train(max_epochs = as.integer(400))
```


## Getting the latent represenation and visualization

```{r}
# get the latent represenation
latent = model$get_latent_representation()

# put it back in our original Seurat object
latent <- as.matrix(latent)
rownames(latent) = colnames(immune.combined)
immune.combined[['scvi']] <- CreateDimReducObject(embeddings = latent, key = "scvi_", assay = DefaultAssay(immune.combined))

```


```{r}
immune.combined <- RunUMAP(immune.combined, dims = 1:20, reduction = 'scvi')
immune.combined <- RunTSNE(immune.combined, dims = 1:20, reduction = 'scvi')
immune.combined <- FindNeighbors(immune.combined, reduction = 'scvi', dims = 1:20) %>% FindClusters(resolution = c(0.2,0.4,0.6,0.8,1.0,1.2,1.4,1.6))
```

# Cell numbers in each resolution

```{r}
statics.cluster.cells <- function(seurat, keys = "(_snn_res)|(^Annotation)"){
  test.data <- as.data.frame(as.matrix(seurat@meta.data[, grepl(keys,colnames(seurat@meta.data))]),stringsAsFactors = FALSE)
  test.data$cell <- rownames(test.data)
  rownames(test.data) <- NULL
  test.data.melt <- reshape2::melt(test.data, id.vars = "cell",value.name = "cluster")
  test.data.melt.dcast <- reshape2::dcast(test.data.melt,formula = cluster ~ variable,fun.aggregate = length)
  test.data.melt.dcast$cluster <- as.numeric(test.data.melt.dcast$cluster)
  test.data.melt.dcast <- arrange(test.data.melt.dcast,cluster)
  rownames(test.data.melt.dcast) <- test.data.melt.dcast$cluster
  test.data.melt.dcast$cluster <- NULL
  return(test.data.melt.dcast)
}
DT::datatable(statics.cluster.cells(immune.combined, keys = paste("(^",DefaultAssay(immune.combined),"_snn_res)|(^Annotation)",sep = "")),  extensions = c('Buttons','ColReorder','FixedColumns'),
              options = list(pageLength = 40, 
                             autoWidth = TRUE,
                             scrollX = TRUE,
                             scrollY = FALSE,
                             dom = 'Bfrtip', # Buttons | note B in dom denotes the place where the buttons are to be inserted
                             buttons = c('csv', 'pdf'),
                             colReorder = TRUE,
                             fixedColumns = TRUE)
              )
```

# Choose Resolution

```{r fig.height=10,fig.width=10}
clustree(immune.combined@meta.data,prefix = paste(DefaultAssay(immune.combined),"_snn_res.",sep = ""))
```

```{r fig.width=8,fig.height=6}
DimPlot(immune.combined,label = TRUE)
DimPlot(immune.combined,label = TRUE,reduction = "tsne")

DimPlot(immune.combined,group.by = "orig.ident")
DimPlot(immune.combined,reduction = "tsne",group.by = "orig.ident")

DimPlot(immune.combined,group.by = "day")
DimPlot(immune.combined,reduction = "tsne",group.by = "day")
```

```{r fig.width=6,fig.height=5}
# DimPlot(immune.combined,group.by = "Phase")
# DimPlot(immune.combined,reduction = "tsne",group.by = "Phase")
```

```{r fig.height=9,fig.width=13}
DimPlot(immune.combined,label = TRUE,split.by = "day",ncol = 4)
```

# Top expressed genes in each cluster

```{r fig.width=6,fig.height=6}
for (i in levels(immune.combined@active.ident)) {
  print(plotHighestExprs(as.SingleCellExperiment(subset(immune.combined,idents = i),assay = DefaultAssay(immune.combined)),colour_cells_by = cluster.resolution))
}
```

```{r}
expr.ave <- AverageExpression(immune.combined, assays = DefaultAssay(immune.combined))[[1]]
expr.ave$gene <- rownames(expr.ave)
expr.ave.melt <- reshape2::melt(expr.ave,id.vars = "gene", value.name = "expr",variable.name = "cluster")
top_20_genes <- expr.ave.melt %>%
  group_by(cluster) %>%
  top_n(20,expr) %>%
  arrange(cluster,desc(expr))

DT::datatable(top_20_genes,  extensions = c('Buttons','ColReorder','FixedColumns'),
              options = list(pageLength = 20, 
                             autoWidth = TRUE,
                             scrollX = FALSE,
                             scrollY = FALSE,
                             dom = 'Bfrtip', # Buttons | note B in dom denotes the place where the buttons are to be inserted
                             buttons = c('csv', 'pdf'),
                             colReorder = FALSE,
                             fixedColumns = FALSE)
              )
```

# Top expressed genes in each time point

```{r fig.width=6,fig.height=6}
Idents(immune.combined) <- "day"
for (i in levels(immune.combined@active.ident)) {
  print(plotHighestExprs(as.SingleCellExperiment(subset(immune.combined,day == i),assay = DefaultAssay(immune.combined)),colour_cells_by = "day"))
}

```

```{r}
expr.ave <- AverageExpression(immune.combined, assays = DefaultAssay(immune.combined))[[1]]
expr.ave$gene <- rownames(expr.ave)
expr.ave.melt <- reshape2::melt(expr.ave,id.vars = "gene", value.name = "expr",variable.name = "cluster")
top_20_genes <- expr.ave.melt %>%
  group_by(cluster) %>%
  top_n(20,expr) %>%
  arrange(cluster,desc(expr))

DT::datatable(top_20_genes,  extensions = c('Buttons','ColReorder','FixedColumns'),
              options = list(pageLength = 20, 
                             autoWidth = TRUE,
                             scrollX = FALSE,
                             scrollY = FALSE,
                             dom = 'Bfrtip', # Buttons | note B in dom denotes the place where the buttons are to be inserted
                             buttons = c('csv', 'pdf'),
                             colReorder = FALSE,
                             fixedColumns = FALSE)
              )
Idents(immune.combined) <- cluster.resolution
```


# Cluster QC

```{r fig.width=20,fig.height=10}
markers <- c("Pdgfra","Epcam","Ptprc","Vwf","Pdgfrb","Nrxn1",'Rgs5','Acta2','Cd34','Il33','Il1rl1','Lgr5')
DefaultAssay(immune.combined) <- "CellBender"
qc.keys <- c(colnames(immune.combined@meta.data)[2:10],"scrublet.doublets.score")
VlnPlot(immune.combined,features = qc.keys, pt.size = 0, ncol = 3) + NoLegend()
VlnPlot(immune.combined,features = markers, pt.size = 0, ncol = 3) + NoLegend()
```

```{r fig.height=12,fig.width=15}
VlnPlot(immune.combined,features = qc.keys, pt.size = 0, ncol = 3, group.by = "day") + NoLegend()
VlnPlot(immune.combined,features = qc.keys, group.by = "day",pt.size = 0)  + NoLegend()
```

```{r fig.width=17,fig.height=12}
# fix reduction selection not working problem in featureplot
cds@reductions$scvitsne@key <- "SCVItSNE_"
colnames(cds@reductions$scvitsne@cell.embeddings) <- paste0("SCVI",colnames(cds@reductions$scvitsne@cell.embeddings))
cds@reductions$scviumap@key <- "SCVIUMAP_"
colnames(cds@reductions$scviumap@cell.embeddings) <- paste0("SCVI", colnames(cds@reductions$scviumap@cell.embeddings))

FeaturePlot(immune.combined,features = markers,reduction = "scviumap") & NoLegend()
```

```{r fig.width=24,fig.height=14}
FeaturePlot(immune.combined,features = qc.keys)
```


# save data

```{r}
if ( !dir.exists(Results.dir)) {
  dir.create(Results.dir,recursive = TRUE)
}

sampleName <- "Merge-All-SeuratHarmony"
saveRDS(immune.combined, file = paste(Results.dir,"/",sampleName,".rds",sep = ""))
```

# save embedings and annotatons to orign rds file

```{r}
# cds <-  readRDS("./4.1-Merged-Rds/Merge-All-SeuratHarmony.rds")
# Reductions(cds)
```

```{r}
# cds[["scvi_umap"]] <- immune.combined@reductions$umap
# DimPlot(cds, reduction = "scvi_umap", pt.size = 0.5)
# cds[["scvi_tsne"]] <- immune.combined@reductions$tsne
# DimPlot(cds, reduction = "scvi_tsne", pt.size = 0.5)
# cds[["scvi"]] <- immune.combined@reductions$scvi
# DimPlot(cds, reduction = "scvi", pt.size = 0.5)
# cds@meta.data[,paste("scvi",grep("^CellBender",colnames(immune.combined@meta.data),value = TRUE),sep = "_")] <- 
#   immune.combined@meta.data[,grep("^CellBender",colnames(immune.combined@meta.data),value = TRUE)]
# # project each dim to genes
# cds <- ProjectDim(cds, reduction = "scvi",dims.print = 1:10)
# saveRDS(cds, file = "./4.1-Merged-Rds/Merge-All-SeuratHarmony.rds")
```





