---
title: "Step2.2 CellBender QC And Generate Raw Cells Rds"
author: "Zhang Yongchao"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: TRUE
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  collapse = TRUE,
  fig.align = "center")
```


```{r}
Results.dir <- "./2.2-Raw-Rds/"
CellRanger.outputs.dir <- "../../../../CellRanger-Outputs/"
```

** cellbender outputs QC And Save Rds **

1. shared cells.

2. Most frequently revised genes.

3. genes with top number of cells in which the UMI counts are revised to 0.

4. Genes with top summed variation in all cells.

```{r}
library(Seurat)
library(Matrix)
library(VennDiagram)
library(dplyr)
```

```{r echo=FALSE}
ReadCB_h5 <- function(filename, use.names = TRUE, unique.features = TRUE) {
  if (!requireNamespace('hdf5r', quietly = TRUE)) {
    stop("Please install hdf5r to read HDF5 files")
  }
  if (!file.exists(filename)) {
    stop("File not found")
  }
  infile <- hdf5r::H5File$new(filename = filename, mode = 'r')
  genomes <- names(x = infile)
  output <- list()
  if (hdf5r::existsGroup(infile, 'matrix')) {
    # cellranger version 3
    message('CellRanger version 3+ format H5')
    if (use.names) {
      feature_slot <- 'features/name'
    } else {
      feature_slot <- 'features/id'
    }
  } else {
    message('CellRanger version 2 format H5')
    if (use.names) {
      feature_slot <- 'gene_names'
    } else {
      feature_slot <- 'genes'
    }
  }
  for (genome in genomes) {
    counts <- infile[[paste0(genome, '/data')]]
    indices <- infile[[paste0(genome, '/indices')]]
    indptr <- infile[[paste0(genome, '/indptr')]]
    shp <- infile[[paste0(genome, '/shape')]]
    features <- infile[[paste0(genome, '/', feature_slot)]][]
    barcodes <- infile[[paste0(genome, '/barcodes')]]
    sparse.mat <- sparseMatrix(
      i = indices[] + 1,
      p = indptr[],
      x = as.numeric(x = counts[]),
      dims = shp[],
      giveCsparse = FALSE
    )
    if (unique.features) {
      features <- make.unique(names = features)
    }
    rownames(x = sparse.mat) <- features
    colnames(x = sparse.mat) <- barcodes[]
    sparse.mat <- as(object = sparse.mat, Class = 'dgCMatrix')
    # Split v3 multimodal
    if (infile$exists(name = paste0(genome, '/features'))) {
      types <- infile[[paste0(genome, '/features/feature_type')]][]
      types.unique <- unique(x = types)
      if (length(x = types.unique) > 1) {
        message("Genome ", genome, " has multiple modalities, returning a list of matrices for this genome")
        sparse.mat <- sapply(
          X = types.unique,
          FUN = function(x) {
            return(sparse.mat[which(x = types == x), ])
          },
          simplify = FALSE,
          USE.NAMES = TRUE
        )
      }
    }
    output[[genome]] <- sparse.mat
  }
  infile$close_all()
  if (length(x = output) == 1) {
    return(output[[genome]])
  } else{
    return(output)
  }
}
```

# import data and create seurat object

```{r}
cellbender.outputs <- paste(CellRanger.outputs.dir,basename(getwd()),"/outs/cellbender.output_filtered.h5",sep = "")
project.name <- basename(getwd())
cellbender_mat <- ReadCB_h5(cellbender.outputs)
colnames(cellbender_mat) <- gsub("-1$","",colnames(cellbender_mat))
cds.cellbender <- CreateSeuratObject(counts = cellbender_mat, project = project.name)
cds.cellbender
```


```{r}
cellranger.outputs <- paste(CellRanger.outputs.dir,basename(getwd()),"/outs/filtered_feature_bc_matrix",sep = "")
project.name <- basename(getwd())
cellranger_mat <- Read10X(cellranger.outputs)
cds.cellranger <- CreateSeuratObject(counts = cellranger_mat, project = project.name)
cds.cellranger
```

# Check Cellbender and CellRanger results

## check both have same features (rownames)

```{r}
all(rownames(cds.cellbender) == rownames(cds.cellranger))
tail(sort(cds.cellbender@meta.data$nCount_RNA,decreasing = TRUE),100)
```

## Q1: what is the zero counts cells

```{r}
cell.zero <- colnames(cds.cellbender)[cds.cellbender@meta.data$nCount_RNA == 0]
table(cell.zero %in% colnames(cds.cellranger))
```

those zero counts cells not exist in cellranger outputs.

## Q2: What is the extra cells in cellbender

```{r}
head(cds.cellbender@meta.data[!colnames(cds.cellbender) %in% colnames(cds.cellranger),])
```

usually cells with low UMI counts and features.

## Q3: Overlapped genes

```{r fig.width=6,fig.height=6}
T<-venn.diagram(list(CellBender=colnames(cds.cellbender),CellRanger=colnames(cds.cellranger)),filename=NULL
                ,lwd=1,lty=2
                ,col=c('red','green'),fill=c('red','green')
                ,cat.col=c('red','green')
                ,rotation.degree=0)
grid.draw(T)
```

## Q4: What is the lost cells in CellRanger

```{r}
cds.cellranger@meta.data[!colnames(cds.cellranger) %in% colnames(cds.cellbender),]
```

Usually cells with low UMI counts!


# UMI Counts Difference

## cell level

```{r fig.width=6,fig.height=6}
shared.cells <- intersect(colnames(cds.cellbender),colnames(cds.cellranger))
y = cds.cellbender@meta.data[shared.cells,"nCount_RNA"]
x = cds.cellranger@meta.data[shared.cells,"nCount_RNA"]
hist(x - y,xlim = c(0,max(x - y)),breaks = max(x - y))
hist((x-y)/x, breaks = 50)
plot(log10(x),x-y)
```

Is dead cells with low UMI counts and doulets with high UMI counts?

## feature level

```{r}
mat.bender <- cellbender_mat[,shared.cells]
mat.ranger <- cellranger_mat[,shared.cells]
mat.dff <- as.matrix(mat.ranger - mat.bender)
```

### changed times

```{r}
change.times <- sort(apply(mat.dff > 0,1,sum),decreasing = TRUE)
plot(1:length(change.times),log(change.times + 1))
head(change.times,30) # cell numbers
```

### changed sum

```{r}
change.sum <- sort(apply(mat.dff,1,sum),decreasing = TRUE)
plot(1:length(change.sum),log(change.sum + 1))
head(change.sum,30) # UMIs
```

### changed mean

```{r}
change.mean <- sort(change.sum/change.times,decreasing = TRUE)
plot(1:length(change.mean),change.mean )
head(change.mean,30)
# mean of genes with top 30 change sum
change.mean[names(head(change.sum,30))]
# mean of genes with top 30 change time
change.mean[names(head(change.times,30))]
```

# save the shared cells

>[Seurat3.0对象及其操作列表 - 多个assay的转换与操作](https://www.jianshu.com/p/b9078c71f057)

```{r}
cds <- subset(cds.cellranger,cells = shared.cells)
cds$nCounts_Diff <-  (x - y)[match(colnames(cds),shared.cells)]
cds$nCounts_Diff_percent <- ((x - y)[match(colnames(cds),shared.cells)])/x
cds[["CellBender"]] <- CreateAssayObject(counts = cellbender_mat[,shared.cells])


if ( !dir.exists(Results.dir)) {
  dir.create(Results.dir)
}

saveRDS(cds,file = paste(Results.dir,basename(getwd()),".rds",sep = ""))
write.csv(as.data.frame(change.sum),file = paste(Results.dir,"Genes-with-summed-revised-UMI-counts.csv",sep = ""))
write.csv(as.data.frame(change.times),file = paste(Results.dir,"Genes-with-summed-revised-cells.csv",sep = ""))
```


# Session info

```{r}
sessionInfo()
```