---
title: "seurat-doubletfinder"
author: "Zhang Yongchao"
date: "February 23, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(DoubletFinder)
```


```{r}
cds <- readRDS("../../PH3-CD326.rds")
DimPlot(cds,label = TRUE)
```

```{r}
sweep.res.list <- paramSweep_v3(cds, PCs = 1:10, sct = FALSE)
sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)
bcmvn <- find.pK(sweep.stats)
```
```{r}
# pK
pK <- as.numeric(as.character(bcmvn$pK[which.max(bcmvn$BCmetric)]))
# pN
library(dplyr)
bc.filtered <- filter(sweep.stats,pK == bcmvn$pK[which.max(bcmvn$BCmetric)])
pN <- as.numeric(as.character(bc.filtered$pN[which.max(bc.filtered$BCreal)]))
print(pK)
print(pN)
```

> [How to determine the number of real doublet #28](https://github.com/chris-mcginnis-ucsf/DoubletFinder/issues/28)
With all of that being said, the Scrublet authors show that their method is amenable to real doublet number estimates. So I would advise testing out their Scrublet if you want to have more confidence in this threshold. Notably, you could use the Scrublet-defined threshold in a DoubletFinder workflow.

>[Which value is the cell loading density value?]()
According to the 10x protocol, this means you should expect about a 3.5% doublet rate.


```{r}
annotations <- cds@meta.data$RNA_snn_res.0.5
homotypic.prop <- modelHomotypic(annotations)           ## ex: annotations <- cds@meta.data$ClusteringResults
nExp_poi <- round(0.035*nrow(cds@meta.data))  ## Assuming 3.5% doublet formation rate - tailor for your dataset
# Scrublet-defined threshold in a DoubletFinder workflow
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
nExp_poi;nExp_poi.adj
```
```{r}
if (nExp_poi != nExp_poi.adj) {
  ## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
  cds <- doubletFinder_v3(cds, PCs = 1:10, pN = pN, pK = pK, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)
  cds <- doubletFinder_v3(cds, PCs = 1:10, pN = pN, pK = pK, nExp = nExp_poi.adj, reuse.pANN = grep("^pANN",colnames(cds@meta.data),value = TRUE), sct = FALSE)
}else{
  cds <- doubletFinder_v3(cds, PCs = 1:10, pN = pN, pK = pK, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)
}
```

```{r}
if (nExp_poi != nExp_poi.adj) {
  DimPlot(cds,group.by = grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1])
  DimPlot(cds,group.by = grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[2])
  cds[[grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1]]] <- factor(cds@meta.data[,grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1]])
  cds[[grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[2]]] <- factor(cds@meta.data[,grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[2]])
  VlnPlot(cds,features = c("nCount_RNA","nFeature_RNA","percent.mt","UMI.Non.mt"),ncol = 4,group.by = grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1],pt.size = 0.01)
  VlnPlot(cds,features = c("nCount_RNA","nFeature_RNA","percent.mt","UMI.Non.mt"),ncol = 4,group.by = grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[2],pt.size = 0.01)
}else{
  DimPlot(cds,group.by = grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1])
  cds[[grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1]]] <- factor(cds@meta.data[,grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1]])
  VlnPlot(cds,features = c("nCount_RNA","nFeature_RNA","percent.mt","UMI.Non.mt"),ncol = 4,group.by = grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1],pt.size = 0.01)
}
```
```{r fig.width=20,fig.height=5}
FeaturePlot(cds,features = c("nCount_RNA", "nFeature_RNA", "percent.mt", "UMI.Non.mt"),ncol = 4,label = TRUE)
```

```{r fig.width=15,fig.height=5}
p1 <- DimPlot(cds,label = TRUE)
p2 <- DimPlot(cds,label = TRUE,cells = rownames(cds@meta.data[cds@meta.data[,grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1]] == "Singlet",]))
p3 <- DimPlot(cds,label = TRUE,cells = rownames(cds@meta.data[cds@meta.data[,grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1]] != "Singlet",]))
p1 + p2 + p3
```
```{r fig.width=15,fig.height=5}
p1 <- DimPlot(cds,label = TRUE,reduction = "tsne")
p2 <- DimPlot(cds,label = TRUE,cells = rownames(cds@meta.data[cds@meta.data[,grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1]] == "Singlet",]),reduction = "tsne")
p3 <- DimPlot(cds,label = TRUE,cells = rownames(cds@meta.data[cds@meta.data[,grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)[1]] != "Singlet",]),reduction = "tsne")
p1 + p2 + p3
```
