---
title: "Step3.2-Remove-Doublets-By-DoublerFinder"
author: "Zhang Yongchao"
date: "March 1, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(DoubletFinder)
library(dplyr)
library(DT)
```

> [How to determine the number of real doublet #28](https://github.com/chris-mcginnis-ucsf/DoubletFinder/issues/28)
With all of that being said, the Scrublet authors show that their method is amenable to real doublet number estimates. So I would advise testing out their Scrublet if you want to have more confidence in this threshold. Notably, you could use the Scrublet-defined threshold in a DoubletFinder workflow.

>[Which value is the cell loading density value?]()
According to the 10x protocol, this means you should expect about a 3.5% doublet rate.

```{r}
# rds.files <- grep("rds$",list.files(),value = TRUE)
```

```{r}
# obj.list <- lapply(rds.files, function(pbmc){
#   message("* Processing sample: ",pbmc)
#   cds <- readRDS(pbmc)
#   return(cds)
# })
# obj.list
```

```{r}
Doublet_Annotation <- function(cds,cluster = "RNA_snn_res.0.6"){
  sweep.res.list <- paramSweep_v3(cds, PCs = 1:20, sct = FALSE)
  sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)
  bcmvn <- find.pK(sweep.stats)
  # pK
  pK <- as.numeric(as.character(bcmvn$pK[which.max(bcmvn$BCmetric)]))
  # pN
  bc.filtered <- filter(sweep.stats,pK == bcmvn$pK[which.max(bcmvn$BCmetric)])
  pN <- as.numeric(as.character(bc.filtered$pN[which.max(bc.filtered$BCreal)]))
  print(paste("The Chosen pK:",pK,"pN:",pN,sep = " "))
  
  # doublet ration
  data <- data.frame(cell.numbers = c(1000,2000,3000,4000,5000),doublets.ratio = c(0.8,1.6,2.3,3.1,3.9))
  # plot(data$cell.numbers,data$doublets.ratio)
  fit <- lm(doublets.ratio~cell.numbers,data)
  doublet.rate <- unname(predict(fit,newdata = data.frame(cell.numbers = 5000)))/100
  print(paste("Doublet Ration is set to:",doublet.rate,sep = " "))
  
  homotypic.prop <- modelHomotypic(cds@meta.data[,cluster])          ## ex: annotations <- cds@meta.data$ClusteringResults
  nExp_poi <- round(doublet.rate*nrow(cds@meta.data)) ## Assuming 3.5% doublet formation rate - tailor for your dataset
  # Scrublet-defined threshold in a DoubletFinder workflow
  nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
  message("Doublet/Adjusted cells:",nExp_poi,";",nExp_poi.adj)
  
  if (nExp_poi != nExp_poi.adj) {
    ## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
    cds <- doubletFinder_v3(cds, PCs = 1:20, pN = pN, pK = pK, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)
    cds <- doubletFinder_v3(cds, PCs = 1:20, pN = pN, pK = pK, nExp = nExp_poi.adj, reuse.pANN = grep("^pANN",colnames(cds@meta.data),value = TRUE), sct = FALSE)
  }else{
    cds <- doubletFinder_v3(cds, PCs = 1:20, pN = pN, pK = pK, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)
  }
  return(cds)
}
```

```{r}
# obj.list <- lapply(obj.list,function(x){
#   cds <- Doublet_Annotation(x)
#   return(cds)
# })
```

# QC for DoubletFinder

```{r}
rds.files <- grep("annotated.rds$",list.files(),value = TRUE)
obj.list <- lapply(rds.files, function(pbmc){
  message("* Processing sample: ",pbmc)
  cds <- readRDS(pbmc)
  return(cds)
})
obj.list
```
## doublet cell numbers

```{r}
statics.doublets <- lapply(obj.list,function(x){
  DF.class <- grep("^DF.classifications",colnames(x@meta.data),value = TRUE)
  if (length(DF.class) == 2) {
    statis.res <- data.frame( sample = x@project.name,
                              doublets = sum(x@meta.data[,DF.class[1]] == "Doublet"),
                              doublets.adjusted = sum(x@meta.data[,DF.class[2]] == "Doublet"))
  }else{
     statis.res <- data.frame( sample = x@project.name,
                              doublets = sum(x@meta.data[,DF.class[1]] == "Doublet"),
                              doublets.adjusted = sum(x@meta.data[,DF.class[1]] == "Doublet"))
  }
  return(statis.res)
})
datatable(Reduce(rbind,statics.doublets))
```

## show doublets in plots

```{r fig.width=6,fig.height=5}
lapply(obj.list,function(cds){
  print(cds@project.name)
  DF.class <- grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)
  if (length(DF.class) == 2) {
    print(DimPlot(cds,group.by = DF.class[1]))
    print(DimPlot(cds,group.by = DF.class[2]))
  }else{
    print(DimPlot(cds,group.by = DF.class[1]))
  }
   return(NULL)
  }
)
```
```{r fig.width=15,fig.height=5}
plot_function <- function(cds,group = "cluster",reduction = "umap"){
  p1 <- DimPlot(cds,label = TRUE,reduction = reduction)
  p2 <- DimPlot(cds,label = TRUE,cells = rownames(cds@meta.data[cds@meta.data[,group] == "Singlet",]),reduction = reduction)
  p3 <- DimPlot(cds,label = TRUE,cells = rownames(cds@meta.data[cds@meta.data[,group] != "Singlet",]),reduction = reduction)
  print(p1 + p2 + p3)
}

lapply(obj.list,function(cds){
  print(cds@project.name)
  DF.class <- grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)
  if (length(DF.class) == 2) {
    plot_function(cds,group = DF.class[1])
    plot_function(cds,group = DF.class[2])
    plot_function(cds,group = DF.class[1],reduction = "tsne")
    plot_function(cds,group = DF.class[2],reduction = "tsne")
  }else{
    plot_function(cds,group = DF.class[1])
    plot_function(cds,group = DF.class[1],reduction = "tsne")
  }
  return(NULL)
  }
)
```



```{r fig.width=20,fig.height=5}
lapply(obj.list,function(cds){
  print(cds@project.name)
  DF.class <- grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)
  if (length(DF.class) == 2) {
    cds[[DF.class[1]]] <- factor(cds@meta.data[,DF.class[1]])
    cds[[DF.class[2]]] <- factor(cds@meta.data[,DF.class[2]])
    print(VlnPlot(cds,features = c("nCount_RNA","nFeature_RNA","percent.mt","UMI.Non.mt"),ncol = 4,group.by = DF.class[1],pt.size = 0.01))
    print(VlnPlot(cds,features = c("nCount_RNA","nFeature_RNA","percent.mt","UMI.Non.mt"),ncol = 4,group.by = DF.class[2],pt.size = 0.01))
  }else{
    cds[[DF.class[1]]] <- factor(cds@meta.data[,DF.class[1]])
    print(VlnPlot(cds,features = c("nCount_RNA","nFeature_RNA","percent.mt","UMI.Non.mt"),ncol = 4,group.by = DF.class[1],pt.size = 0.01))
  }
  return(NULL)
  }
)
```

```{r}
# lapply(obj.list, function(x){
#   print(paste("*Processing sample: ",x@project.name,sep = ""))
#   sampleName <- paste(x@project.name,".doublets.annotated.rds",sep = "")
#   saveRDS(x,file = sampleName)
#   return(NULL)
# })
```

```{r fig.width=12,fig.height=12}
par(mfrow = c(4,4))
lapply(obj.list,function(cds){
  print(cds@project.name)
  pANN <- grep("^pANN",colnames(cds@meta.data),value = TRUE)
  DF.class <- grep("^DF.classifications",colnames(cds@meta.data),value = TRUE)
  if (length(DF.class) == 2) {
    print(hist(cds@meta.data[,pANN],breaks = 100,xlab = "PANN score",ylab = "cell counts",main = cds@project.name))
    print(raw.cut <- min(cds@meta.data[cds@meta.data[,DF.class[1]] == "Doublet",pANN]))
    print(adj.cut <- min(cds@meta.data[cds@meta.data[,DF.class[2]] == "Doublet",pANN]))
    print(abline(v=c(raw.cut,adj.cut),lwd = 2,col = c("blue","red")))
  }else{
    print(hist(cds@meta.data[,pANN],breaks = 100,xlab = "PANN score",ylab = "cell counts",main = cds@project.name))
    print(raw.cut <- min(cds@meta.data[cds@meta.data[,DF.class[1]] == "Doublet",pANN]))
    print(abline(v=raw.cut,lwd = 2,col = "blue"))
  }
  return(NULL)
  }
)
```








