---
title: "Step 4.1 Merge all samples by RPCA"
author: "Zhang Yongchao"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: TRUE
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      warning = FALSE)
```

Method Refer:

>[Seurat](https://satijalab.org/seurat/)

>[Fast integration using reciprocal PCA (RPCA)](https://satijalab.org/seurat/articles/integration_rpca.html)

**Parameters:**

```{r}
Results.dir <- "./4.1-Merged-Rds/RPCA-Method"
```


```{r}
suppressMessages(library(dplyr)) 
suppressMessages(library(Seurat))
suppressMessages(library(patchwork))
suppressMessages(library(cowplot))
suppressMessages(library(ggplot2))
suppressMessages(library(ggpubr))
suppressMessages(library(clustree))
suppressMessages(library(harmony))
suppressMessages(library(scater))
suppressMessages(library(DT))
```

# import and merge all samples

```{r}
AnalysisDirs <- list.dirs("../../SingleSample/Cellbender-Seurat-Scrublet",recursive = FALSE,full.names = TRUE)
rds.files <- list.files(paste(AnalysisDirs,"/3.3-Clean-Cells-Rds",sep = ""),pattern = "rds$",full.names = TRUE)
basename(rds.files)
```


```{r}
obj.list <- lapply(rds.files, function(pbmc){
  # message("* Processing sample: ",pbmc)
  cds <- readRDS(pbmc)
  # message("Project Name is ", cds@project.name)
  return(cds)
})
# obj.list
```

```{r}
# remove unwanted sample: P_LI_DAdult_B1
# obj.list <- obj.list[-15]
```

# Perform integration and Relevel all samples

```{r}
features <- SelectIntegrationFeatures(object.list = obj.list)
```

```{r}
immune.anchors <- FindIntegrationAnchors(object.list = obj.list, anchor.features = features, reduction = "rpca")
immune.combined <- IntegrateData(anchorset = immune.anchors)

# factor the cluster after merge datas
columns.factorization <- function(ADataFrame,columns = c(1:2)){
  for (i in columns) {
    ADataFrame[,i] <- as.factor(as.numeric(ADataFrame[,i]))
  }
  return(ADataFrame)
}

immune.combined@meta.data <- columns.factorization(immune.combined@meta.data,columns = grep("_snn_res",colnames(immune.combined@meta.data),value = TRUE))
```

```{r}
immune.combined@meta.data$day <- unlist(lapply(strsplit(immune.combined@meta.data$orig.ident, split = "_"),function(x){paste(x[c(1,3)],collapse = "_")}))
immune.combined@meta.data$stage <- ifelse(grepl("^E",immune.combined@meta.data$orig.ident),"Embryo","PostNatal")

immune.combined@meta.data$orig.ident <- factor(immune.combined@meta.data$orig.ident, 
                                               levels = c("E_SI_D115_B1", "E_SI_D115_B2","E_SI_D125_B1", "E_SI_D125_B2", 
                                                          "E_SI_D135_B1", "E_SI_D135_B2", "E_SI_D135_B3", "E_SI_D145_B0", "E_SI_D155_B0",
                                                          "E_SI_D195_B0", "E_SI_D195_B1", "E_SI_D195_B2", "P_SI_D1_B0", "P_SI_D7_B0", "P_SI_D21_B0",
                                                          "P_SI_DAdult_B0") )
immune.combined@meta.data$day <- factor(immune.combined@meta.data$day,levels = c("E_D115", "E_D125", "E_D135", "E_D145", "E_D155", "E_D195", "P_D1","P_D7","P_D21","P_DAdult"))
```

# Filter Merged Samples

```{r}
# immune.combined <- subset(immune.combined, subset = nCount_RNA < 1e5 & nFeature_RNA > 700 & percent.mt < 20 & nCount_RNA > 3000)
# immune.combined <- subset(immune.combined, percent.mt < 15 & nCount_RNA < 8e4)

# Attention! do not use code bellow when run new data.
# immune.combined <- subset(immune.combined, scrublet < 0.1)
```

# Merged QC - Post Filter

```{r}
# Code as Same as Pre filter
```


# Add Sample Annotations and cells numbers

```{r}
# immune.combined@meta.data$sex <- ifelse(as.character(immune.combined@meta.data$orig.ident) %in% c("E_LI_D115_B1", "E_LI_D115_B2", "P_LI_D1_B1", "P_LI_D1_B2", "P_LI_D21_B1"), "Most Male","Male and Female")
# table(immune.combined@meta.data$sex)

# Cell numbers by each annotation
table(immune.combined@meta.data$orig.ident)
table(immune.combined@meta.data$day)
table(immune.combined@meta.data$stage)
```

# Modify old cluster names - Prefixed with Single

```{r}
colnames(immune.combined@meta.data)[grepl("(S.Score)|(G2M.Score)|(Phase)|(_snn_res)|(^Annotation)",colnames(immune.combined@meta.data))] <- 
  paste("Single_",colnames(immune.combined@meta.data)[grepl("(S.Score)|(G2M.Score)|(Phase)|(_snn_res)|(^Annotation)",colnames(immune.combined@meta.data))],sep = "")
```

## Perform an integrated analysis

```{r fig.width=18,fig.height=7}
DefaultAssay(immune.combined) <- "integrated"
```

```{r  fig.height=6,fig.width=7}
immune.combined <- ScaleData(immune.combined, verbose = FALSE)
immune.combined <- RunPCA(immune.combined, npcs = 30, verbose = FALSE, assay = DefaultAssay(immune.combined))
```

```{r fig.height=5,fig.width=6}
DimPlot(immune.combined, reduction = "pca")
```

```{r fig.height=25,fig.width=15}
DimHeatmap(immune.combined, dims = 1:15, cells = 500, balanced = TRUE)
```

```{r fig.width=6,fig.height=5}
ElbowPlot(immune.combined)
```


```{r}
immune.combined <- RunUMAP(immune.combined, dims = 1:30,umap.method = "umap-learn",metric = "correlation", assay = DefaultAssay(immune.combined))
immune.combined <- RunTSNE(immune.combined, dims = 1:30, assay = DefaultAssay(immune.combined))
immune.combined <- FindNeighbors(immune.combined, dims = 1:30, assay = DefaultAssay(immune.combined)) %>% FindClusters(resolution = c(0.2,0.4,0.6,0.8,1.0,1.2,1.4,1.6), assay = DefaultAssay(immune.combined))
```


# Cell numbers in each resolution

```{r}
statics.cluster.cells <- function(seurat, keys = "(_snn_res)|(^Annotation)"){
  test.data <- as.data.frame(as.matrix(seurat@meta.data[, grepl(keys,colnames(seurat@meta.data))]),stringsAsFactors = FALSE)
  test.data$cell <- rownames(test.data)
  rownames(test.data) <- NULL
  test.data.melt <- reshape2::melt(test.data, id.vars = "cell",value.name = "cluster")
  test.data.melt.dcast <- reshape2::dcast(test.data.melt,formula = cluster ~ variable,fun.aggregate = length)
  test.data.melt.dcast$cluster <- as.numeric(test.data.melt.dcast$cluster)
  test.data.melt.dcast <- arrange(test.data.melt.dcast,cluster)
  rownames(test.data.melt.dcast) <- test.data.melt.dcast$cluster
  test.data.melt.dcast$cluster <- NULL
  return(test.data.melt.dcast)
}
DT::datatable(statics.cluster.cells(immune.combined, keys = paste("(^",DefaultAssay(immune.combined),"_snn_res)|(^Annotation)",sep = "")),  extensions = c('Buttons','ColReorder','FixedColumns'),
              options = list(pageLength = 40, 
                             autoWidth = TRUE,
                             scrollX = TRUE,
                             scrollY = FALSE,
                             dom = 'Bfrtip', # Buttons | note B in dom denotes the place where the buttons are to be inserted
                             buttons = c('csv', 'pdf'),
                             colReorder = TRUE,
                             fixedColumns = TRUE)
              )
```

# Choose Resolution

```{r fig.height=10,fig.width=10}
clustree(immune.combined@meta.data,prefix = paste(DefaultAssay(immune.combined),"_snn_res.",sep = ""))
```

```{r}
cluster.resolution <- paste(DefaultAssay(immune.combined),"_snn_res.0.6",sep = "")
# cluster.resolution <- "RNA_snn_res.0.6"
Idents(immune.combined) <- cluster.resolution
```

```{r fig.width=8,fig.height=6}
DimPlot(immune.combined,label = TRUE)
DimPlot(immune.combined,label = TRUE,reduction = "tsne")

DimPlot(immune.combined,group.by = "orig.ident")
DimPlot(immune.combined,reduction = "tsne",group.by = "orig.ident")

DimPlot(immune.combined,group.by = "day")
DimPlot(immune.combined,reduction = "tsne",group.by = "day")
```


```{r fig.height=9,fig.width=13}
DimPlot(immune.combined,label = TRUE,split.by = "day",ncol = 4)
```

# Top expressed genes in each cluster

```{r fig.width=6,fig.height=6}
# DefaultAssay(immune.combined) <- "CellBender"
# 
# for (i in levels(immune.combined@active.ident)) {
#   print(plotHighestExprs(as.SingleCellExperiment(subset(immune.combined,idents = i),assay = DefaultAssay(immune.combined)),colour_cells_by = cluster.resolution))
# }
```

```{r}
# expr.ave <- AverageExpression(immune.combined, assays = DefaultAssay(immune.combined))[[1]]
# expr.ave$gene <- rownames(expr.ave)
# expr.ave.melt <- reshape2::melt(expr.ave,id.vars = "gene", value.name = "expr",variable.name = "cluster")
# top_20_genes <- expr.ave.melt %>%
#   group_by(cluster) %>%
#   top_n(20,expr) %>%
#   arrange(cluster,desc(expr))
# 
# DT::datatable(top_20_genes,  extensions = c('Buttons','ColReorder','FixedColumns'),
#               options = list(pageLength = 20, 
#                              autoWidth = TRUE,
#                              scrollX = FALSE,
#                              scrollY = FALSE,
#                              dom = 'Bfrtip', # Buttons | note B in dom denotes the place where the buttons are to be inserted
#                              buttons = c('csv', 'pdf'),
#                              colReorder = FALSE,
#                              fixedColumns = FALSE)
#               )
```

# Top expressed genes in each time point

```{r fig.width=6,fig.height=6}
# Idents(immune.combined) <- "day"
# for (i in levels(immune.combined@active.ident)) {
#   print(plotHighestExprs(as.SingleCellExperiment(subset(immune.combined,day == i),assay = DefaultAssay(immune.combined)),colour_cells_by = "day"))
# }

```

```{r}
# expr.ave <- AverageExpression(immune.combined, assays = DefaultAssay(immune.combined))[[1]]
# expr.ave$gene <- rownames(expr.ave)
# expr.ave.melt <- reshape2::melt(expr.ave,id.vars = "gene", value.name = "expr",variable.name = "cluster")
# top_20_genes <- expr.ave.melt %>%
#   group_by(cluster) %>%
#   top_n(20,expr) %>%
#   arrange(cluster,desc(expr))
# 
# DT::datatable(top_20_genes,  extensions = c('Buttons','ColReorder','FixedColumns'),
#               options = list(pageLength = 20, 
#                              autoWidth = TRUE,
#                              scrollX = FALSE,
#                              scrollY = FALSE,
#                              dom = 'Bfrtip', # Buttons | note B in dom denotes the place where the buttons are to be inserted
#                              buttons = c('csv', 'pdf'),
#                              colReorder = FALSE,
#                              fixedColumns = FALSE)
#               )
# Idents(immune.combined) <- cluster.resolution
```


# Cluster QC

```{r fig.width=20,fig.height=10}
markers <- c("Pdgfra","Epcam","Ptprc","Vwf","Pdgfrb","Nrxn1",'Rgs5','Acta2','Cd34','Il33','Il1rl1','Lgr5')
qc.keys <- c(colnames(immune.combined@meta.data)[2:10],"scrublet.doublets.score")

VlnPlot(immune.combined,features = qc.keys, pt.size = 0, ncol = 3) + NoLegend()
VlnPlot(immune.combined,features = markers, pt.size = 0, ncol = 3) + NoLegend()
```

```{r fig.height=12,fig.width=15}
VlnPlot(immune.combined,features = qc.keys, pt.size = 0, ncol = 3, group.by = "day") + NoLegend()
VlnPlot(immune.combined,features = qc.keys, group.by = "day",pt.size = 0)  + NoLegend()
```

```{r fig.width=17,fig.height=12}
FeaturePlot(immune.combined,features = markers) & NoLegend()
```

```{r fig.width=24,fig.height=14}
FeaturePlot(immune.combined,features = qc.keys)
```


# save data

```{r}
if ( !dir.exists(Results.dir)) {
  dir.create(Results.dir,recursive = TRUE)
}

sampleName <- "Merge-All-SeuratHarmony"
saveRDS(immune.combined, file = paste(Results.dir,"/",sampleName,".rds",sep = ""))
```

