---
title: "Step3.5 Automated Cluster Annotation"
author: "Zhang Yongchao"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: TRUE
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  collapse = TRUE,
  fig.align = "center")
```

```{r}
Clean.rds.dir <- "./3.3-Clean-Cells-Rds"
Results.SCSA.dir <- "./3.5-Cluster-Annotation/SCSA"
Cluster.Markers.dir <- "./3.4-Downstream-Analysis/ClusterMarkers"
SCSA.py.dir <- "/home/xilab/software/Python-SCSA/SCSA/"
```

```{r}
suppressMessages(library(dplyr)) 
suppressMessages(library(Seurat))
suppressMessages(library(patchwork))
suppressMessages(library(ggplot2))
suppressMessages(library(clustree))
suppressMessages(library(SingleR))
suppressMessages(library(pheatmap))
suppressMessages(library(DT))
suppressMessages(library(scCATCH))
```

# import data and set parameters

```{r}
rds.file <- list.files(Clean.rds.dir,pattern = ".rds$",full.names = TRUE)
cds <- readRDS(rds.file)
cds
```

**Parameters:**

```{r}
cluster.resolution <- paste(DefaultAssay(cds),"_snn_res.0.6",sep = "")
# cluster.resolution <- "RNA_snn_res.0.6"
Idents(cds) <- cluster.resolution
```

# SingleR

> [Using SingleR to annotate single-cell RNA-seq data](https://bioconductor.org/packages/release/bioc/vignettes/SingleR/inst/doc/SingleR.html)

```{r}
hpca.se <- MouseRNAseqData()
hpca.se
table(hpca.se$label.main)
```

## performed on single cells 

```{r}
umi.matrix <- cds[[DefaultAssay(cds)]]@data
pred.hesc <- SingleR(test = umi.matrix, ref = hpca.se, labels = hpca.se$label.main)
head(pred.hesc)
table(pred.hesc$first.labels)
sum(is.na(pred.hesc$pruned.labels))
```

```{r fig.width=30,fig.height=8}
plotScoreHeatmap(pred.hesc,annotation_col = cds@meta.data[,c(paste(DefaultAssay(cds),"_snn_res.0.2",sep = ""), cluster.resolution)])
```

```{r}
all(rownames(pred.hesc) == rownames(cds@meta.data))
table(pred.hesc$first.labels,cds@meta.data[,cluster.resolution])
tab <- table(Assigned=pred.hesc$pruned.labels, Cluster=cds@active.ident)
# Adding a pseudo-count of 10 to avoid strong color jumps with just 1 cell.
pheatmap(log2(tab+10), color=colorRampPalette(c("white", "blue"))(101))

table(pred.hesc$first.labels,cds@meta.data[,paste(DefaultAssay(cds),"_snn_res.0.2",sep = "")])
tab <- table(Assigned=pred.hesc$pruned.labels, Cluster=cds@meta.data[,paste(DefaultAssay(cds),"_snn_res.0.2",sep = "")])
# Adding a pseudo-count of 10 to avoid strong color jumps with just 1 cell.
pheatmap(log2(tab+10), color=colorRampPalette(c("white", "blue"))(101))
```


```{r fig.height=10,fig.width=8}
plotScoreDistribution(pred.hesc)
```


## aggregated into cluster-level profiles prior to annotation

```{r}
pred.cluster <- SingleR(test = umi.matrix, 
                        ref = hpca.se, 
                        labels = hpca.se$label.main,
                        method = "cluster",
                        clusters = cds@active.ident)
```


```{r fig.width=8,fig.height=4}
plotScoreHeatmap(pred.cluster,show.labels = TRUE)
```
```{r fig.height=10,fig.width=8}
plotScoreDistribution(pred.cluster)
```

## save annotation

```{r fig.width=9,fig.height=6}
cds@meta.data[,"Annotation.SingleR.SingleCell.level"] <- as.factor(pred.hesc$labels)
cds@meta.data[,"Annotation.SingleR.Cluster.level"] <- as.factor(pred.cluster$labels[match(as.character(cds@active.ident),rownames(pred.cluster))])
DimPlot(cds,label = TRUE,group.by = "Annotation.SingleR.SingleCell.level")
DimPlot(cds,label = TRUE,group.by = "Annotation.SingleR.Cluster.level")
```

# SCSA

> [SCSA: cell type annotation for single-cell RNA-seq data(https://github.com/bioinfo-ibms-pumc/SCSA)

## code

```{r}
if ( !dir.exists(Results.SCSA.dir)) {
  dir.create(Results.SCSA.dir,recursive = TRUE)
}

SCSA.py.path <- paste(SCSA.py.dir,"SCSA.py",sep = "")
SCSA.db.path <- paste(SCSA.py.dir,"whole.db",sep = "")
Markers.csv <- list.files(Cluster.Markers.dir,full.names = TRUE)[1]
Results.txt <- paste(Results.SCSA.dir,"pythonSCSA.Results.txt", sep = "/")
Results.log <- paste(Results.SCSA.dir,"pySCSA.log", sep = "/")
```

SCSA code Demo

```{bash eval = FALSE}
# This code will not run for setting: eval = FALSE
python3 /home/xilab/software/Python-SCSA/SCSA/SCSA.py \
-d /home/xilab/zhangyc/Python-SCSA/SCSA/whole.db \
-s seurat \
-i  ./3.4-Downstream-Analysis/DEGs/*.csv \
-k All \
-E \
-g Mouse \
-m txt \
-f 0.25 \
-o ./3.5-Cluster-Annotation/SCSA/pythonSCSA.Results.txt \
>& ./3.5-Cluster-Annotation/SCSA/pySCSA.log
```


```{r}
SCSA.command <- paste("python3",SCSA.py.path,"-d",SCSA.db.path,"-s seurat -i",Markers.csv,"-k All -E -g Mouse -m txt -f 0.25 -o",Results.txt,">&",Results.log)
SCSA.command
```

```{r}
system(SCSA.command)
```

## results - cluster annotation info

```{r}
anno <- read.delim(Results.txt,stringsAsFactors = FALSE)
anno <- anno %>% 
  dplyr::group_by(Cluster) %>%
  top_n(10, Z.score)
DT::datatable(anno ) %>% formatSignif(columns="Z.score", digits=4)
```

## results - cluster GO info

```{r}
GO <- read.delim(paste(Results.txt,".go",sep = ""),stringsAsFactors = FALSE)
GO <- GO[GO$q.value < 0.05,]
DT::datatable(GO) %>% formatSignif(columns=c('p.value', 'q.value'), digits=2)
```

## save Annotation info to rds file

**Attention**: Only use the labels with highest score to annotate clusters

```{r fig.width=9,fig.height=6}
cluster.info <- anno %>% 
  group_by(Cluster) %>%
  top_n(1,Z.score)
cds@meta.data[,"Annotation.SCSA.Cluster.level"] <- as.factor(cluster.info$Cell.Type[match(as.numeric(as.character(cds@active.ident)),cluster.info$Cluster)])
DimPlot(cds,label = TRUE,group.by = "Annotation.SCSA.Cluster.level")
```

# CellAssign

> [cellassign](https://github.com/Irrationone/cellassign)

Not used here, You need provide a makers file from a well annotated scRNAseq data!

# Garnett

> [garnett](https://cole-trapnell-lab.github.io/garnett/)

Not used here, You need provide a well annotated scRNAseq data to train a cell type classifier.

# scCATCH

> [ZJUFanLab/scCATCH](https://github.com/ZJUFanLab/scCATCH)

> Cell Types:

1. [Human](https://github.com/ZJUFanLab/scCATCH/wiki/human_tissues)

2. [Mouse](https://github.com/ZJUFanLab/scCATCH/wiki/mouse_tissues)

I thought the rannotation quality depens on the cell marker reference, while I do not found celltypes of mesen or endothelium...

## code

```{r}
clu_markers <- read.csv(Markers.csv,row.names = 1,stringsAsFactors = FALSE)
clu_ann<- scCATCH(clu_markers, species = 'Mouse', tissue = c("Colon", "Colon epithelium", "Fetal intestine", "Gastrointestinal tract", "Ileum", "Intestinal crypt", "Intestine", "Mesenteric lymph node", "Small intestine",
                                                             "Fetal brain","Embryo", "Embryoid body", "Embryonic heart", "Embryonic stem cell","Fetal stomach"))
```

## results

```{r}
DT::datatable(clu_ann[,-2])
```

## save to rds file

```{r fig.width=9,fig.height=6}
cds@meta.data[,"Annotation.scCATCH.Cluster.level"] <- as.factor(clu_ann$cell_type[match(as.numeric(as.character(cds@active.ident)),clu_ann$cluster)])
DimPlot(cds,label = TRUE,group.by = "Annotation.scCATCH.Cluster.level")
```


# Save Final results

```{r}
saveRDS(cds,file = rds.file)
```

# Session info

```{r}
sessionInfo()
```

