---
title: "Wei Ruxue - negative selection in RKO cell lines using method from 2017 Cancer Research"
author: "Zhang Yongchao"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: TRUE
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

Using Mageck output as input to find the APC SL Metablolism genes.

# Prepare and set specific parameter

Set special parameters for this analysis

```{r}
samplePrefix <- "RKO."
```

import library

```{r}
suppressMessages(library(ggplot2))
suppressMessages(library(psych))
suppressMessages(library(dplyr))
suppressMessages(library(factoextra))
suppressMessages(library(DT))
suppressMessages(library(ggpubr)) 
suppressMessages(library(reshape2)) 
```



# import Mageck output and  QC

## import data

```{r}
MageckOut <- read.delim("../mageck/Metabolism.count.txt",stringsAsFactors = F)
datatable(head(MageckOut),extensions = 'FixedColumns',options = list(scrollY = FALSE,scrollX = TRUE))
# value columns
ValueColumns <- 3:length(colnames(MageckOut))
```


```{r}
### Attention! this code chunk used to modify the data!
## Metabolism Day14/Day0 samples may mistakely labeled!
# colnamesOrder <- colnames(MageckOut)
# colnames(MageckOut)[grep("Metabolism.Day14",colnames(MageckOut))] <- gsub("Metabolism.Day14","tmp",colnames(MageckOut)[grep("Metabolism.Day14",colnames(MageckOut))])
# colnames(MageckOut)[grep("Metabolism.Day0",colnames(MageckOut))] <- gsub("Metabolism.Day0","Metabolism.Day14",colnames(MageckOut)[grep("Metabolism.Day0",colnames(MageckOut))])
# colnames(MageckOut)[grep("tmp",colnames(MageckOut))] <- gsub("tmp","Metabolism.Day0",colnames(MageckOut)[grep("tmp",colnames(MageckOut))])
# print(colnamesOrder)
# print(colnames(MageckOut))
# # reorder columns as old!
# MageckOut <- MageckOut[,colnamesOrder]
# print(colnames(MageckOut))
```

three types of sgRNAs-index

```{r}
# index of essential genes
LibraryMetadata <- read.csv("../../../library/Library2-APC_SL_metabolism+essential+nontargeting.csv",stringsAsFactors = FALSE)
EssentialGenes <- filter(LibraryMetadata,Type == "essential")
EssentialIndex <- which(MageckOut$sgRNA %in% EssentialGenes$ID)
# index of the Non-targeting sgRNA index
NonTargetIndex <- which(substring(MageckOut$Gene,1,5) == "NonTa")
# index of the customed sgRNAs(except for the essential and nontargeting)
CustomedIndex <- seq(1,length(rownames(MageckOut)))[!seq(1,length(rownames(MageckOut))) %in% c(EssentialIndex,NonTargetIndex)]
```

## QC-1: Count reads

### Counts reads from  1% top enriched sgRNAs 


```{r}
TopsgRNAsNumer <- round(length(rownames(MageckOut))*0.01)
TopsgRNAs <- apply(MageckOut[,ValueColumns],2,function(x)MageckOut$sgRNA[order(x,decreasing = TRUE)[1:TopsgRNAsNumer]])
TopsgRNAsGene <- apply(MageckOut[,ValueColumns],2,function(x)MageckOut$Gene[order(x,decreasing = TRUE)[1:TopsgRNAsNumer]])

# top 1% enriched sgRNAs in each sample
datatable(TopsgRNAs,options = list(scrollY = FALSE,scrollX = TRUE))
# top 1% enriched-sgRNAS-Matched Genes in each sample
datatable(TopsgRNAsGene,options = list(scrollY = FALSE,scrollX = TRUE))

topsgRNAsReads <- apply(MageckOut[,ValueColumns],2,function(x)x[order(x,decreasing = TRUE)])[1:TopsgRNAsNumer,]
totalReadsSums <- apply(MageckOut[,ValueColumns], 2, sum)
topsgRNAsSums <- apply(topsgRNAsReads, 2, sum)
topsgRNAsFre <- round(topsgRNAsSums/totalReadsSums,3)

CountsResults <- rbind(totalReadsSums,topsgRNAsSums,topsgRNAsFre)
```

### counts sgRNA reads from essential genes/Non-targeting sgRNAs/Customed genes

```{r}
indexes <- c("NonTargetIndex","EssentialIndex","CustomedIndex")
VectorTailChange <- function(somevector,TailPosition,newValue) {
  somevector[(length(somevector)-TailPosition + 1):length(somevector)] <- newValue
  return(somevector)
}

for (index in indexes) {
  typeSumName <- gsub("Index","-sgRNAS-Sum",index)
  typeFreName <- gsub("Index","-sgRNAS-Sum-Fre",index)
  assign(typeSumName,apply(MageckOut[get(index),ValueColumns],2,sum))
  assign(typeFreName,round(get(typeSumName)/totalReadsSums,3))
  CountsResults <- rbind(CountsResults,get(typeSumName),get(typeFreName))
  rownames(CountsResults) <- VectorTailChange(rownames(CountsResults),2,c(typeSumName,typeFreName))
}
datatable(CountsResults,extensions = 'FixedColumns',options = list(scrollY = FALSE,scrollX = TRUE,fixedColumns = TRUE))
```




## QC-2: Distribution describe

```{r}
datatable(describe(MageckOut),extensions = 'FixedColumns',options = list(scrollY = FALSE,scrollX = TRUE,fixedColumns = TRUE))
```

All sgRNAs are found with more than 1 reads!

## QC-3: Visualize reads distribution in histogram

Raw read counts distribution in a histogram

```{r,  fig10, fig.width=12,fig.height=16}
par(mfrow = c(4,3))
for (i in ValueColumns) {
  hist(MageckOut[,i],breaks = seq(0,300000,by=500),xlim = c(0,20000), main = colnames(MageckOut)[i],xlab = NULL)
  rug(jitter(MageckOut[,i]))
}
```

```{r fig21,fig.width=9,fig.height=36}
par(mfrow = c(12,3))
for (i in ValueColumns) {
  for (index in indexes) {
    hist(MageckOut[get(index),i],breaks = seq(0,300000,by=500),xlim = c(0,25000), main = paste(colnames(MageckOut)[i],index,sep = "-"),xlab = NULL)
    rug(jitter(MageckOut[,i]))
  }
}
```

# Normalize and Calculate the FC, then normalized to the median nontargeting control sgRNAs

## Normalize library size and QC

### Normalized to same total reads

```{r}
# normalized to total read count per sample (normalized reads per sgRNA = (reads per sgRNA/total reads for all sgRNAs in the sample) * 1e6 + 1)
MageckNorm <- MageckOut
MageckNorm[,3:length(colnames(MageckNorm))] <- apply(MageckOut[,ValueColumns],2,function(x)((x/sum(x))*1e6 + 1))
colSums(MageckNorm[,3:length(colnames(MageckNorm))])# each sample is normalized to 1001606 reads! 
```

###  QC-1: Boxplot 

Normalized read count distribution of all samples in a boxplot (normalized to total read count per sample)

```{r fig1, fig.height=8,fig.width=10}
par(oma=c(5,5,5,5))
boxplot(MageckNorm[,ValueColumns], main = "sgRNA abundance",col = c(rep("#C6DBEF",3), rep("#2171B5",3), rep("#FDD0A2",3), rep("#D94801",3)), pch = 21,cex = 0.6,ylim = c(0,2000), xaxt = 'n',ylab = "Normalized reads counts")
axis(side = 1, at = seq(1,12), labels = gsub(samplePrefix,"",colnames(MageckOut)[ValueColumns]),las = "2")

# log2(Normalized counts - total 1001606 reads)
boxplot(log2(MageckNorm[,ValueColumns]), main = "sgRNA abundance",col = c(rep("#C6DBEF",3), rep("#2171B5",3), rep("#FDD0A2",3), rep("#D94801",3)), pch = 21,cex = 0.6,xaxt = 'n',ylab = "log2(Normalized reads counts)")
axis(side = 1, at = seq(1,12), labels = gsub(samplePrefix,"",colnames(MageckOut)[ValueColumns]),las = "2")
```

### QC-2: Scatter plot with Rsquare

```{r fig2, fig.width=12, fig.height=16}
# using Normalized counts (total about 1e6 counts)
rsquare <- data.frame(sample1=character(0),sample2=character(0),rsquare=character(0),stringsAsFactors = FALSE)
# Suppress one command's output in R [here is legend function]
# https://stackoverflow.com/questions/2723034/suppress-one-commands-output-in-r
hush=function(code){
  sink("/dev/null") # use /dev/null in UNIX
  tmp = code
  sink()
  return(tmp)
}

 for (i in ValueColumns) {
  print(paste("#### Comparison of ",colnames(MageckOut)[i],"with all other samples using Normalized counts ####",sep = " "))
  par(mfrow = c(4,3))
  hush(for (j in ValueColumns) {
    sampleA <- colnames(MageckOut)[i]
    sampleB <- colnames(MageckOut)[j]
    #compName <- paste(sampleA,sampleB,sep = " vs ")
    print(plot(MageckNorm[,c(i,j)],pch = 16,cex=0.5),xlim = c(0,6500),ylim=c(0,7000))
    fit <- summary(lm(MageckNorm[,j] ~ MageckNorm[,i]))
    legend("bottomright", bty="n", legend=paste("R2 = ", format(fit$adj.r.squared, digits=4)))
    newEntry <- data.frame(sample1=sampleA,sample2=sampleB,rsquare=format(fit$adj.r.squared, digits=4),stringsAsFactors = FALSE)
    rsquare <- rbind(rsquare,newEntry)
    #rm(sampleA,sampleB,fit,newEntry,compName)
  })
}
```

```{r}
print("#### You can check the R square for each comparison using Normalized counts in the table bellow! ####")
datatable(rsquare)
```

```{r fig3, fig.width=12, fig.height=16}
# using log2(Normalized counts)
rsquare <- data.frame(sample1=character(0),sample2=character(0),rsquare=character(0),stringsAsFactors = FALSE)
log2MageckNorm <- log2(MageckNorm[,ValueColumns])
for (i in 1:(length(colnames(MageckOut))-2)) {
  print(paste("#### Comparison of ",colnames(MageckOut)[i+2],"with all other samples using log2(Normalized counts ####)",sep = " "))
  par(mfrow = c(4,3))
  hush(for (j in 1:(length(colnames(MageckOut))-2)) {
    sampleA <- colnames(MageckOut)[i+2]
    sampleB <- colnames(MageckOut)[j+2]
    #compName <- paste(sampleA,sampleB,sep = " vs ")
    print(plot(log2MageckNorm[,c(i,j)],pch = 16,cex=0.5))
    fit <- summary(lm(log2MageckNorm[,j] ~ log2MageckNorm[,i]))
    print(legend("bottomright", bty="n", legend=paste("R2 = ", format(fit$adj.r.squared, digits=4))))
    newEntry <- data.frame(sample1=sampleA,sample2=sampleB,rsquare=format(fit$adj.r.squared, digits=4),stringsAsFactors = FALSE)
    rsquare <- rbind(rsquare,newEntry)
    #rm(sampleA,sampleB,fit,newEntry,compName)
  })
}
```

```{r}
print("#### You can check the R square for each comparison using log2(Normalized counts) in the table bellow! ####")
datatable(rsquare)
```

### QC-3: PCA

http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/

```{r}
# using normlized data
MageckPCA <- MageckNorm[,ValueColumns]
rownames(MageckPCA) <- MageckNorm$sgRNA
MageckPCA <- t(MageckPCA)
res.pca <- prcomp(MageckPCA, scale = TRUE)
fviz_eig(res.pca)
fviz_pca_ind(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
)

fviz_pca_ind(res.pca,
             col.ind = c(rep("WT.Day0",3),rep("Metabolism.Day0",3),rep("WT.Day14",3),rep("Metabolism.Day14",3)),
             repel = TRUE     # Avoid text overlapping
) +
  theme(legend.position="none")
```

```{r}
# using log2(normlized data)
MageckPCA <- log2(MageckPCA)
res.pca <- prcomp(MageckPCA, scale = TRUE)
fviz_eig(res.pca)
fviz_pca_ind(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
)

fviz_pca_ind(res.pca,
             col.ind = c(rep("red",3),rep("green",3),rep("blue",3),rep("black",3)),
             repel = TRUE     # Avoid text overlapping
) +
  theme(legend.position="none")
```

## Calculate log(FC) And QC

Calculate log(FC) for each pair (Day14 vs Day0)

### Calculate log(FC) for each Day14-vs-Day0

```{r}
Replicates <- paste("rep",1:3,sep = "")
Pairs <- c( "WT","Metabolism")
# Compares <- c("Day14","Day0") # not used here

extracname <- function(object,x,y,z="."){
  tmp <- object
  for (i in c(x, y, z)) {
  tmp <- grep(i, tmp, value = TRUE)
  }
  return(tmp)
}
  
object <- colnames(MageckNorm)
#extracname(object,"WT","Day14","rep1")

attach(MageckNorm)
for (Replicate in Replicates) {
  for(Pair in Pairs){
    compareName <- paste(Pair,"Day14.vs.Day0",Replicate,"logFC",sep = ".")
    MageckNorm$tmp <- log(get(extracname(object,Replicate,Pair,"Day14"))/get(extracname(object,Replicate,Pair,"Day0")))
    colnames(MageckNorm)[which(colnames(MageckNorm) == "tmp")]  <- compareName
    rm(compareName)
  }
}
detach(MageckNorm)
datatable(head(MageckNorm), extensions = 'FixedColumns',options = list(scrollY = FALSE,scrollX = TRUE,fixedColumns = list(leftColumns = 2)))
```

### QC: Log(FC) Histogram

Log(FC) Value distribution before normalized to the median of the nontargeting control sgRNAs for each sample.

#### Total reads

```{r fig11,fig.width=8,fig.height=12}
par(mfrow=c(3,2))
for (name in grep("logFC",colnames(MageckNorm),value = T)) {
  hist(MageckNorm[,name],breaks = seq(-5,5,by = 0.05), main = name, xlab = NULL,xlim = c(-3,2))
}

```

#### Three types of reads

```{r, fig14,fig.width=8,fig.height=16}
simplifyName <- function(ObjectName,Removekeys,Newkeys){
  for (i in 1:length(Removekeys)) {
    ObjectName <- gsub(Removekeys[i],Newkeys[i],ObjectName)
  }
  return(ObjectName)
}
#simplifyName("WT.Day14.vs.Day0.rep1.logFC",c("Day14.vs.Day0.","logFC"),c("","Non-targeting"))
oldNameKeys <- c("Day14.vs.Day0.","logFC")

par(mfrow=c(6,3))
for (name in grep("logFC",colnames(MageckNorm),value = T)) {
  for (index in indexes) {
    hist(MageckNorm[get(index),name],breaks = seq(-2.5,2,by = 0.1), main = simplifyName(name,oldNameKeys,c("",index)), xlab = NULL,xlim = c(-1.5,1))
  }
}
```

## Normalized log(FC) and QC

### Normalize logFC to the median of the nontargeting control sgRNAs

```{r}
NonTargetMean <- as.numeric(apply(MageckNorm[NonTargetIndex,grep("logFC",colnames(MageckNorm))],2,median)) # Using median value as the scale factor
print("Non-target sgRNA median:")
print(NonTargetMean)
MageckNorm[,grep("logFC",colnames(MageckNorm))] <- sweep(MageckNorm[,grep("logFC",colnames(MageckNorm))],2,NonTargetMean,"-")
colnames(MageckNorm)[grep("logFC",colnames(MageckNorm))] <- paste(grep("logFC",colnames(MageckNorm),value = T),".Normalized",sep = "")
write.csv(MageckNorm,file = "Day14-vs-Day0-Normalized-logFC.csv")
```


### QC: Value distribution in Histogram

#### Total reads
```{r fig9,fig.width=8,fig.height=12}
par(mfrow=c(3,2))
for (name in grep("logFC.Normalized",colnames(MageckNorm),value = T)) {
  hist(MageckNorm[,name],breaks = seq(-5,5,by = 0.05), main = name, xlab = NULL,xlim = c(-1.5,1.5))
}
```

#### Three type of reads
```{r fig19,fig.width=8,fig.height=16}
par(mfrow=c(6,3))
oldNameKeys <- c("Day14.vs.Day0.","logFC")

par(mfrow=c(6,3))
for (name in grep("logFC",colnames(MageckNorm),value = T)) {
  for (index in indexes) {
      hist(MageckNorm[get(index),name],breaks = seq(-2.5,2,by = 0.1), main = simplifyName(name,oldNameKeys,c("",index)), xlab = NULL,xlim = c(-1,1))
  }
}
```

### QC: Value distribution in Boxplot

```{r fig4, fig.height=7,fig.width=7}
par(oma=c(4,1,1,1))
boxplot(MageckNorm[,grep("logFC.Normalized",colnames(MageckNorm))], main = "Value Distribution",col = c(rep("#C6DBEF",3), rep("#2171B5",3)), pch = 21,cex = 0.6, xaxt = 'n',ylab = "Normalized log(FC) - Day14.vs.Day0",ylim = c(-2,1))
axis(side = 1, at = seq(1,6), labels = simplifyName(grep("logFC.Normalized",colnames(MageckNorm),value = TRUE),c(".logFC.Normalized",".Day14.vs.Day0"),c("","")),las = "2")
```



# Visualize genes/sgRNAs in a scatter plot

```{r,fig12,fig.width=12,fig.height=12}
# parameters for ggplot2
limitation <- c(-4,4)
breaks <- c(-4,-3,-2,-1,0,1,2,3,4)

for (i in Replicates) {
  print(paste("####################################****",i,"****####################################",sep=""))
  # extract paired data for each replicate
  data <- MageckNorm[,extracname(colnames(MageckNorm),"logFC",i)]
  attach(data)
  caseName <- grep("Metabolism",colnames(data),value = TRUE)
  controlName <- grep("WT",colnames(data),value = TRUE)
  colnames(data)[grep("Metabolism",colnames(data))] <- "case"
  colnames(data)[grep("WT",colnames(data))] <- "control"
  
  # Total sgRNAs
  p1 <-ggplot(data, aes(x=control, y=case)) +
    geom_vline(xintercept = 0) + geom_hline(yintercept = 0) +
    geom_point(size = 0.3,color = "#bed2ed") +
    lims(x=limitation,y=limitation) +
    theme_minimal() +
    coord_fixed() +
    scale_x_continuous(breaks = breaks,limits = limitation) +
    scale_y_continuous(breaks = breaks,limits = limitation) +
    labs(x=controlName,y=caseName,title="Total sgRNAs") 
  #print(p1)
  
  # Highlight NonTargeting sgRNAs
  p2 <-ggplot(data, aes(x=control, y=case)) +
    geom_vline(xintercept = 0) + geom_hline(yintercept = 0) +
    geom_point(size = 0.3,color = "#bed2ed") +
    #lims(x=c(-6,6),y=c(-6,6)) +
    geom_point(data = data[NonTargetIndex,],aes(x=control, y=case),size = 0.3,color = "#7c7f7e") +
    theme_minimal() +
    coord_fixed() +
    scale_x_continuous(breaks = breaks,limits = limitation) +
    scale_y_continuous(breaks = breaks,limits = limitation) +
    labs(x=controlName,y=caseName,title="Highlight NonTargeting sgRNAs") 
  #print(p2)
  
  # highlight essential genes/sgRNAs
  p3 <-ggplot(data, aes(x=control, y=case)) +
    geom_vline(xintercept = 0) + geom_hline(yintercept = 0) +
    geom_point(size = 0.2,color = "#bed2ed") +
    geom_point(data = data[EssentialIndex,],aes(x=control, y=case),size = 0.2,color = "#ed2144") +
    #lims(x=c(-10,10),y=c(-10,10)) +
    theme_minimal() +
    coord_fixed()  +
    scale_x_continuous(breaks = breaks,limits = limitation) +
    scale_y_continuous(breaks = breaks,limits = limitation) +
    labs(x=controlName,y=caseName,title="Highlight essential sgRNAs") 
  #print(p3)
  
  # highligth customed sgRNAs
  p5 <-ggplot(data, aes(x=control, y=case)) +
    geom_vline(xintercept = 0) + geom_hline(yintercept = 0) +
    geom_point(size = 0.2,color = "#bed2ed") +
    geom_point(data = data[CustomedIndex,],aes(x=control, y=case),size = 0.2,color = "#ed2144") +
    #lims(x=c(-10,10),y=c(-10,10)) +
    theme_minimal() +
    coord_fixed()   +
    scale_x_continuous(breaks = breaks,limits = limitation) +
    scale_y_continuous(breaks = breaks,limits = limitation) +
    labs(x=controlName,y=caseName,title="Highlight customed sgRNAs") 
  #print(p5)
  
  print(ggarrange(p1,p2,p3,p5,labels = c("A","B","C","D"),nrow = 2,ncol = 2))
  
  # # check sgRNA counts
  # geneName <- "BTNL9"
  # geneIndex <- which(MageckNorm$Gene %in% geneName)

  # p7 <-ggplot(data, aes(x=control, y=case)) +
  #   geom_vline(xintercept = 0) + geom_hline(yintercept = 0) +
  #   geom_point(size = 0.2,color = "#bed2ed") +
  #   geom_point(data = data[geneIndex,],aes(x=control, y=case),size = 0.2,color = "red") +
  #   #lims(x=c(-10,10),y=c(-10,10)) +
  #   theme_minimal() +
  #   coord_fixed()   +
  #   scale_x_continuous(breaks = breaks,limits = limitation) +
  #   scale_y_continuous(breaks = breaks,limits = limitation) +
  #   labs(x=controlName,y=caseName,title=paste("Highlight gRNAs from",gneName,sep = " ")) 
  # p7
  detach(data)
}

```




# filter sgRNAs using customed standards

use non-targeting sgRNAs to define intervals of c("positive","non-essential","negative")

```{r}
# first use non-targeting sgRNAs to set cutoffs for the WT and APC-mutant
quantile.cut <- c(0.95, 0.95) # 0.95 for WT, 1 for APC-mutant
limitation <- c(-1,1)
breaks <- c(-2,-1.5,-1,-0.5,-0.2,-0.1,0,0.1,0.2,0.5,1,1.5,2)

for (i in Replicates) {
  print(paste("####################################****",i,"****####################################",sep=""))
  # extract paired data for each replicate
  data <- MageckNorm[,c("sgRNA","Gene",extracname(colnames(MageckNorm),"logFC",i))]
  attach(data)
  caseName <- grep("Metabolism",colnames(data),value = TRUE)
  controlName <- grep("WT",colnames(data),value = TRUE)
  colnames(data)[grep("Metabolism",colnames(data))] <- "case"
  colnames(data)[grep("WT",colnames(data))] <- "control"
  nonTarget <- data[NonTargetIndex,c("control","case")]
  distance.cut <- sqrt(apply(apply(nonTarget,2,function(x)(x*x)),2,function(x)quantile(x, quantile.cut)))
  print(distance.cut)
  data.filtered <- filter(data,abs(control) < distance.cut[1,1] & case <  distance.cut[2,2]*(-1))
  print("Distance to origin Cutoff:")
  print(paste(controlName,distance.cut[1,1],sep = ": "))
  print(paste(caseName,distance.cut[2,2],sep = ": "))
  datatable(data.filtered)
  data.filtered$Replicate <- i
  data.filtered$ControlCutoff <- distance.cut[1,1]
  data.filtered$CaseCutoff <- distance.cut[2,2]
  # merge the selected sgRNAs
  if (i == Replicates[1]) {
    merged.selection <- data.filtered
  }else{
    merged.selection <- rbind(merged.selection,data.filtered)
  }
  
  # check sgRNA position!
  sgRNAIndex <- which(MageckNorm$sgRNA %in% data.filtered$sgRNA)
  p7 <-ggplot(data, aes(x=control, y=case)) +
    geom_vline(xintercept = 0) + geom_hline(yintercept = 0) +
    geom_hline(yintercept = -distance.cut[2,2],colour = "black",linetype = "dotted") + 
    geom_vline(xintercept = c(-distance.cut[1,1],distance.cut[1,1]),colour = "black",linetype = "dotted") +
    geom_point(size = 0.2,color = "#bed2ed") +
    geom_point(data = data[sgRNAIndex,],aes(x=control, y=case),size = 0.2,color = "red") +
    theme_minimal() +
    coord_fixed()   +
    scale_x_continuous(breaks = breaks,limits = limitation) +
    scale_y_continuous(breaks = breaks,limits = limitation) +
    labs(x=controlName,y=caseName,title="Highlight the selected gRNAs") 
  print(p7)
  
  detach(data)
}
stats.Gene <- dcast(as.data.frame(table(merged.selection$Gene,merged.selection$Replicate)),Var1~Var2)
rownames(stats.Gene) <- stats.Gene$Var1
stats.Gene <- stats.Gene[,-1]
stats.Gene$sum <- apply(stats.Gene,1,sum)
datatable(stats.Gene)
```




