---
title: "Step 4.1 Merge Replicates by Seurat Harmony"
author: "Zhang Yongchao"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: TRUE
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      warning = FALSE)
```

Method Refer:

>[Seurat](https://satijalab.org/seurat/)

>[SeuratWrappers](https://github.com/satijalab/seurat-wrappers)

>[Integration of datasets using Harmony](http://htmlpreview.github.io/?https://github.com/satijalab/seurat-wrappers/blob/master/docs/harmony.html)

**Parameters:**

```{r}
Results.dir <- "./4.1-Replicates-Merged-Cells-Rds"
Single.dir <- "../../SingleSample/"
Single.Rds.dir <- "/3.3-Clean-Cells-Rds"
sampleName <- basename(getwd())
QC.markers <- c("Pdgfra","Epcam","Ptprc","Vwf","Pdgfrb","Nrxn1",'Rgs5','Acta2','Cd34','Il33','Il1rl1','Lgr5')
```


```{r}
suppressMessages(library(dplyr)) 
suppressMessages(library(Seurat))
suppressMessages(library(patchwork))
suppressMessages(library(cowplot))
suppressMessages(library(ggplot2))
suppressMessages(library(ggpubr))
suppressMessages(library(clustree))
suppressMessages(library(harmony))
suppressMessages(library(scater))
suppressMessages(library(DT))
```

# import and merge all samples

```{r}
AnalysisDirs <- list.dirs(Single.dir,recursive = FALSE,full.names = TRUE)
AnalysisDirs <- AnalysisDirs[gsub("_B\\d+.*$","",basename(AnalysisDirs)) == basename(getwd())]
if (any(grepl("Merge$",basename(AnalysisDirs)))) {
  # Only keep merged rds
  keys <- gsub("Merge","",grep("Merge$",basename(AnalysisDirs), value = TRUE))
  sample.OneLane <- AnalysisDirs[grep(keys,basename(AnalysisDirs),invert = TRUE)]
  AnalysisDirs <- c(sample.OneLane,AnalysisDirs[grepl("Merge$",basename(AnalysisDirs))])
}

rds.files <- list.files(paste(AnalysisDirs,Single.Rds.dir,sep = ""),pattern = "rds$",full.names = TRUE)
basename(rds.files)
```

```{r}
obj.list <- lapply(rds.files, function(pbmc){
  # message("* Processing sample: ",pbmc)
  cds <- readRDS(pbmc)
  # message("Project Name is ", cds@project.name)
  return(cds)
})
# obj.list
```

# Merge and Relevel all samples

```{r}
immune.combined <- merge(obj.list[[1]], obj.list[-1])
DefaultAssay(immune.combined) <- "RNA"

# factor the cluster after merge datas
columns.factorization <- function(ADataFrame,columns = c(1:2)){
  for (i in columns) {
    ADataFrame[,i] <- as.factor(as.numeric(ADataFrame[,i]))
  }
  return(ADataFrame)
}

immune.combined@meta.data <- columns.factorization(immune.combined@meta.data,columns = grep("_snn_res",colnames(immune.combined@meta.data),value = TRUE))
```

```{r}
immune.combined@meta.data$day <- unlist(lapply(strsplit(immune.combined@meta.data$orig.ident, split = "_"),function(x){paste(x[c(1,3)],collapse = "_")}))
```


# Merged QC - Pre Filter

```{r fig.width=24,fig.height=20}
qc.keys <- c(colnames(immune.combined@meta.data)[2:6],"scrublet.doublets.score")
VlnPlot(immune.combined, features = qc.keys, ncol = 4,group.by = "orig.ident",pt.size = 0.01)
VlnPlot(immune.combined, features = qc.keys, ncol = 4,group.by = "orig.ident",pt.size = 0)
```


```{r fig.width=15,fig.height=10}
metadata <- immune.combined@meta.data
  
# Visualize the number UMIs/transcripts per cell
p1 <- metadata %>% 
  	ggplot(aes(color=orig.ident, x=nCount_RNA, fill= orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
    geom_vline(xintercept = c(1500,2000,3000,5e4,7e4,1e5)) + NoLegend()

p2 <- metadata %>% 
  	ggplot(aes(color=orig.ident, x=nCount_RNA_RemoveMT, fill= orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
    geom_vline(xintercept = c(1500,2000,3000,5e4,7e4,1e5))

# Visualize the distribution of features detected per cell
p3 <- metadata %>% 
  	ggplot(aes(color=orig.ident, x=nFeature_RNA, fill=orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = c(300,400,500,700,800)) + NoLegend()


# Visualize the distribution of mitochondrial gene expression detected per cell
p4 <- metadata %>% 
  	ggplot(aes(color=orig.ident, x=percent.mt, fill=orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = c(20,30,40,50,60,70))


# Visualize the overall complexity of the gene expression by visualizing the genes detected per UMI
p5 <- metadata %>%
  	ggplot(aes(x=log10GenesPerUMI, color = orig.ident, fill=orig.ident)) +
  	geom_density(alpha = 0.2) +
  	theme_classic() +
  	geom_vline(xintercept = c(0.7,0.75,0.8,0.85)) + NoLegend()

wrap_plots(p1,p2,p3,
           p4,p5)
```


```{r fig.width=12,fig.height=8}
# Visualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIs
x.line <- c(1000,2000,3000,5e4,1e5)
y.line <- c(300,500,700)

p1 <- metadata %>% 
  	ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = x.line) +
  	geom_hline(yintercept = y.line)

p2 <- metadata %>% 
  	ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=log10GenesPerUMI)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = x.line) +
  	geom_hline(yintercept = y.line)

p3 <- metadata %>% 
  	ggplot(aes(x=nCount_RNA_RemoveMT, y=nFeature_RNA, color=percent.mt)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = x.line) +
  	geom_hline(yintercept = y.line)

p4 <- metadata %>% 
  	ggplot(aes(x=nCount_RNA_RemoveMT, y=nFeature_RNA, color=log10GenesPerUMI)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = x.line) +
  	geom_hline(yintercept = y.line)


wrap_plots(p1,p2,
           p3,p4)
```


```{r fig.width=16,fig.height=8}
plot1 <- FeatureScatter(immune.combined, feature1 = paste("nCount", DefaultAssay(immune.combined),sep = "_"), feature2 = "percent.mt",group.by = "orig.ident") + NoLegend()
plot2 <- FeatureScatter(immune.combined, feature1 = paste("nCount", DefaultAssay(immune.combined),sep = "_"), feature2 = paste("nFeature", DefaultAssay(immune.combined),sep = "_"), group.by = "orig.ident")
plot1 + plot2

plot1 <- FeatureScatter(immune.combined, feature1 = paste("nCount", DefaultAssay(immune.combined),"RemoveMT",sep = "_"), feature2 = "percent.mt", group.by = "orig.ident") + NoLegend()
plot2 <- FeatureScatter(immune.combined, feature1 = paste("nCount", DefaultAssay(immune.combined),"RemoveMT",sep = "_"), feature2 = paste("nFeature", DefaultAssay(immune.combined),sep = "_"), group.by = "orig.ident")
plot1 + plot2
```

```{r fig.width=6,fig.height=6}
plotHighestExprs(as.SingleCellExperiment(immune.combined,assay = DefaultAssay(immune.combined)),colour_cells_by = "orig.ident")
```

# Filter Merged Samples

```{r}
# immune.combined <- subset(immune.combined, subset = nCount_RNA < 1e5 & nFeature_RNA > 700 & percent.mt < 20 & nCount_RNA > 3000)
# immune.combined <- subset(immune.combined, percent.mt < 15 & nCount_RNA < 8e4)

# Attention! do not use code bellow when run new data.
# immune.combined <- subset(immune.combined, scrublet < 0.1)
```

# Merged QC - Post Filter

```{r}
# Code as Same as Pre filter
```

# Check sample sex info

```{r fig.width=8,fig.height=4}
VlnPlot(immune.combined,features = c("Xist"),group.by = "orig.ident",pt.size = 0)
```

# Add Sample Annotations and cells numbers

```{r}
# immune.combined@meta.data$sex <- ifelse(as.character(immune.combined@meta.data$orig.ident) %in% c("E_LI_D115_B1", "E_LI_D115_B2", "P_LI_D1_B1", "P_LI_D1_B2", "P_LI_D21_B1"), "Most Male","Male and Female")
# table(immune.combined@meta.data$sex)

# Cell numbers by each annotation
table(immune.combined@meta.data$orig.ident)
table(immune.combined@meta.data$day)
```

# Modify old cluster names - Prefixed with Single

```{r}
colnames(immune.combined@meta.data)[grepl("(S.Score)|(G2M.Score)|(Phase)|(_snn_res)|(^Annotation)",colnames(immune.combined@meta.data))] <- 
  paste("Single_",colnames(immune.combined@meta.data)[grepl("(S.Score)|(G2M.Score)|(Phase)|(_snn_res)|(^Annotation)",colnames(immune.combined@meta.data))],sep = "")
```

## Perform an integrated analysis

```{r fig.width=18,fig.height=7}
s.genes <- readRDS("~/reference/Biology-Database/Cell-Cycles-Genes/Mouse/mouse.s.genes.rds")
g2m.genes <- readRDS("~/reference/Biology-Database/Cell-Cycles-Genes/Mouse/mouse.g2m.genes.rds")
# s.genes
s.genes
# g2m.genes
g2m.genes


immune.combined <- NormalizeData(immune.combined, assay = DefaultAssay(immune.combined))
immune.combined <- FindVariableFeatures(immune.combined, selection.method = "vst", nfeatures = 4000, assay = DefaultAssay(immune.combined))
# Remove mito genes from variable gnes
grep("^mt-",VariableFeatures(immune.combined),value = TRUE)
VariableFeatures(immune.combined) <- grep("^mt-",VariableFeatures(immune.combined),value = TRUE,invert = TRUE)
# Remove Cell Cycle-Related Genes
VariableFeatures(immune.combined) <- VariableFeatures(immune.combined)[! VariableFeatures(immune.combined) %in% c(s.genes,g2m.genes)]
length(VariableFeatures(immune.combined))

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(immune.combined), 10)
# plot variable features with and without labels
plot1 <- VariableFeaturePlot(immune.combined)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

```{r  fig.height=6,fig.width=7}
immune.combined <- CellCycleScoring(immune.combined, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE, assay = DefaultAssay(immune.combined))
immune.combined <- ScaleData(immune.combined, vars.to.regress = c("S.Score", "G2M.Score","percent.mt",paste("nCount", DefaultAssay(immune.combined),sep = "_")), 
                             features = VariableFeatures(immune.combined), verbose = FALSE, assay = DefaultAssay(immune.combined))

immune.combined <- RunPCA(immune.combined, features = VariableFeatures(immune.combined), npcs = 30, verbose = FALSE, assay = DefaultAssay(immune.combined))
# Examine and visualize PCA results a few different ways
print(immune.combined[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(immune.combined, dims = 1:2, reduction = "pca")
```

```{r fig.height=5,fig.width=6}
DimPlot(immune.combined, reduction = "pca")
```

```{r fig.height=25,fig.width=15}
DimHeatmap(immune.combined, dims = 1:15, cells = 500, balanced = TRUE)
```

```{r fig.width=6,fig.height=5}
ElbowPlot(immune.combined)
```

```{r fig.width=6,fig.height=5}
immune.combined <- RunHarmony(immune.combined, group.by.vars = "orig.ident", assay = DefaultAssay(immune.combined))
ElbowPlot(immune.combined,reduction = "harmony")
```

```{r}
immune.combined <- RunUMAP(immune.combined, reduction = "harmony", dims = 1:30, assay = DefaultAssay(immune.combined))
immune.combined <- RunTSNE(immune.combined, reduction = "harmony", dims = 1:30, assay = DefaultAssay(immune.combined))
immune.combined <- FindNeighbors(immune.combined, reduction = "harmony", dims = 1:30, assay = DefaultAssay(immune.combined)) %>% 
  FindClusters(resolution = c(0.2,0.4,0.6,0.8,1.0,1.2,1.4,1.6), assay = DefaultAssay(immune.combined))
```


# Cell numbers in each resolution

```{r}
statics.cluster.cells <- function(seurat, keys = "(_snn_res)|(^Annotation)"){
  test.data <- as.data.frame(as.matrix(seurat@meta.data[, grepl(keys,colnames(seurat@meta.data))]),stringsAsFactors = FALSE)
  test.data$cell <- rownames(test.data)
  rownames(test.data) <- NULL
  test.data.melt <- reshape2::melt(test.data, id.vars = "cell",value.name = "cluster")
  test.data.melt.dcast <- reshape2::dcast(test.data.melt,formula = cluster ~ variable,fun.aggregate = length)
  test.data.melt.dcast$cluster <- as.numeric(test.data.melt.dcast$cluster)
  test.data.melt.dcast <- arrange(test.data.melt.dcast,cluster)
  rownames(test.data.melt.dcast) <- test.data.melt.dcast$cluster
  test.data.melt.dcast$cluster <- NULL
  return(test.data.melt.dcast)
}
DT::datatable(statics.cluster.cells(immune.combined, keys = paste("(^",DefaultAssay(immune.combined),"_snn_res)|(^Annotation)",sep = "")),  extensions = c('Buttons','ColReorder','FixedColumns'),
              options = list(pageLength = 40, 
                             autoWidth = TRUE,
                             scrollX = TRUE,
                             scrollY = FALSE,
                             dom = 'Bfrtip', # Buttons | note B in dom denotes the place where the buttons are to be inserted
                             buttons = c('csv', 'pdf'),
                             colReorder = TRUE,
                             fixedColumns = TRUE)
              )
```

# Choose Resolution

```{r fig.height=10,fig.width=10}
clustree(immune.combined@meta.data,prefix = paste(DefaultAssay(immune.combined),"_snn_res.",sep = ""))
```

```{r}
cluster.resolution <- paste(DefaultAssay(immune.combined),"_snn_res.0.6",sep = "")
# cluster.resolution <- "RNA_snn_res.0.6"
Idents(immune.combined) <- cluster.resolution
```

```{r fig.width=8,fig.height=6}
DimPlot(immune.combined,label = TRUE)
DimPlot(immune.combined,label = TRUE,reduction = "tsne")

DimPlot(immune.combined,group.by = "orig.ident")
DimPlot(immune.combined,reduction = "tsne",group.by = "orig.ident")
```

```{r fig.width=6,fig.height=5}
DimPlot(immune.combined,group.by = "Phase")
DimPlot(immune.combined,reduction = "tsne",group.by = "Phase")
```


# Top expressed genes in each cluster

```{r fig.width=6,fig.height=6}
for (i in levels(immune.combined@active.ident)) {
  print(plotHighestExprs(as.SingleCellExperiment(subset(immune.combined,idents = i),assay = DefaultAssay(immune.combined)),colour_cells_by = cluster.resolution))
}
```

```{r}
expr.ave <- AverageExpression(immune.combined, assays = DefaultAssay(immune.combined))[[1]]
expr.ave$gene <- rownames(expr.ave)
expr.ave.melt <- reshape2::melt(expr.ave,id.vars = "gene", value.name = "expr",variable.name = "cluster")
top_20_genes <- expr.ave.melt %>%
  group_by(cluster) %>%
  top_n(20,expr) %>%
  arrange(cluster,desc(expr))

DT::datatable(top_20_genes,  extensions = c('Buttons','ColReorder','FixedColumns'),
              options = list(pageLength = 20, 
                             autoWidth = TRUE,
                             scrollX = FALSE,
                             scrollY = FALSE,
                             dom = 'Bfrtip', # Buttons | note B in dom denotes the place where the buttons are to be inserted
                             buttons = c('csv', 'pdf'),
                             colReorder = FALSE,
                             fixedColumns = FALSE)
              )
```

# Cluster QC

```{r fig.width=20,fig.height=10}
VlnPlot(immune.combined,features = qc.keys, pt.size = 0, ncol = 3) + NoLegend()
VlnPlot(immune.combined,features = QC.markers, pt.size = 0, ncol = 3) + NoLegend()
```


```{r fig.width=17,fig.height=12}
FeaturePlot(immune.combined,features = QC.markers) & NoLegend()
```

```{r fig.width=24,fig.height=14}
FeaturePlot(immune.combined,features = qc.keys)
```

# Compare Merged and Single Results

```{r fig.height=10,fig.width=10}
p <- DimPlot(immune.combined,label = TRUE,group.by = paste("Single_",DefaultAssay(immune.combined),"_snn_res.0.6",sep = ""),split.by = "orig.ident")
p1 <- DimPlot(obj.list[[1]],label = TRUE,group.by = paste(DefaultAssay(immune.combined),"_snn_res.0.6",sep = ""))
p2 <- DimPlot(obj.list[[2]],label = TRUE,group.by = paste(DefaultAssay(immune.combined),"_snn_res.0.6",sep = ""))

p /
  (p1 | p2)
```


# save data

```{r}
if ( !dir.exists(Results.dir)) {
  dir.create(Results.dir,recursive = TRUE)
}
saveRDS(immune.combined, file = paste(Results.dir,"/",sampleName,".rds",sep = ""))
```

