---
title: "Seurat for xenium data analysis"
author: "Zhang Yongchao"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## preparation

```{r init, message=FALSE, warning=FALSE}
library(Seurat)
library(future)
library(qs2)
plan("multisession", workers = 2)
library(spacexr)
options(future.globals.maxSize= 60*1024^3)
```

```{r}
cropped.obj <- JoinLayers(cropped.obj,assay = 'Xenium')
qs_save(cropped.obj,file = 'featal-SI_SCT_joined.qs2')
cropped.obj <- qs_read('featal-SI_SCT.qs2')
```

```{r rctd.qeury, warning=FALSE}

query.counts <- GetAssayData(cropped.obj, assay = "Xenium", slot = "counts")[, Cells(cropped.obj[["zoom"]])]
coords <- GetTissueCoordinates(cropped.obj[["zoom"]], which = "centroids")
rownames(coords) <- coords$cell
coords$cell <- NULL
query <- SpatialRNA(coords, query.counts, colSums(query.counts))
```

```{r rctd.reference, eval=FALSE}
# allen.corted.ref can be downloaded here: https://www.dropbox.com/s/cuowvm4vrf65pvq/allen_cortex.rds?dl=1
allen.cortex.ref <- readRDS("/data2/shared_data_backup/jiangdx/scRNAseq/scRNAseq-Mouse-Small-Intestine/Analysis/Merge-SI-Embryo-Without-E115/Merge-All-without-E115/4.1-Merged-Rds/Merge-All-Embryo-SeuratHarmony.rds")
allen.cortex.ref <- UpdateSeuratObject(allen.cortex.ref)

Idents(allen.cortex.ref) <- "Annotation.celltype"
# remove CR cells because there aren't enough of them for annotation
# allen.cortex.ref <- subset(allen.cortex.ref, subset = subclass != "CR")
counts <- GetAssayData(allen.cortex.ref, assay = "RNA", slot = "counts")
cluster <- allen.cortex.ref$Annotation.celltype
names(cluster) <- colnames(allen.cortex.ref)
nUMI <- allen.cortex.ref$nCount_RNA
names(nUMI) <- colnames(allen.cortex.ref)
nUMI <- colSums(counts)
# levels(cluster) <- gsub("/", "-", levels(cluster))
reference <- Reference(counts, cluster, nUMI)
```

```{r niche.run.rctd, warning=FALSE, results=FALSE, eval=FALSE}
# run RCTD with many cores
RCTD <- create.RCTD(query, reference, max_cores = 12)
RCTD <- run.RCTD(RCTD, doublet_mode = "doublet")
```

```{r niche.add.annotations, eval=FALSE}
annotations.df <- RCTD@results$results_df
annotations <- annotations.df$first_type
names(annotations) <- rownames(annotations.df)
xenium.obj$predicted.celltype <- annotations
keep.cells <- Cells(xenium.obj)[!is.na(xenium.obj$predicted.celltype)]
xenium.obj <- subset(xenium.obj, cells = keep.cells)
```

While the previous analyses consider each cell independently, spatial data enables cells to be defined not just by their neighborhood, but also by their broader spatial context. In Seurat v5, we introduce support for 'niche' analysis of spatial data, which demarcates regions of tissue ('niches'), each of which is defined by a different composition of spatially adjacent cell types. Inspired by methods in [Goltsev et al, Cell 2018](https://www.sciencedirect.com/science/article/pii/S0092867418309048) and [He et al, NBT 2022](https://www.nature.com/articles/s41587-022-01483-z), we consider the 'local neighborhood' for each cell - consisting of its `k.neighbor` spatially closest neighbors, and count the occurrences of each cell type present in this neighborhood. We then use k-means clustering to group cells that have similar neighborhoods together, into spatial niches.

We call the `BuildNicheAssay` function from within Seurat to construct a new assay called `niche` containing the cell type composition spatially neighboring each cell. A metadata column called `niches` is also returned, which contains cluster assignments based on the niche assay.

```{r build.niche.assay, eval=FALSE}
xenium.obj <- BuildNicheAssay(
  object = xenium.obj,
  fov = "crop",
  group.by = "predicted.celltype",
  niches.k = 5,
  neighbors.k = 30
)
```

```{r load.niche.results, eval=TRUE, include=FALSE}
# xenium.obj <- readRDS("/brahms/hartmana/vignette_data/xenium_niches_presaved.rds")
```

We can then group cells either by their cell type identity, or their niche identity. The niches identified clearly demarcate the neuronal layers in the cortex.

```{r, niche.dimplots, fig.width=8, fig.height=6, warning=FALSE}
celltype.plot <- ImageDimPlot(
  xenium.obj,
  group.by = "predicted.celltype",
  size = 1.5,
  cols = "polychrome",
  dark.background = F) +
  ggtitle("Cell type")
niche.plot <- ImageDimPlot(
  xenium.obj,
  group.by = "niches",
  size = 1.5,
  dark.background = F) +
  ggtitle("Niches") +
  scale_fill_manual(
    values = c("#442288", "#6CA2EA", "#B5D33D", "#FED23F", "#EB7D5B"))
celltype.plot | niche.plot
```

Further, we observe that the composition of each niche is enriched for distinct cell types.

```{r niche.composition}
table(xenium.obj$predicted.celltype, xenium.obj$niches)
```

